Application of Machine Learning to Predict the Optoelectronic Properties of Benzobisazoles

### 2014-06-30 paper/paper.tex
Add updates to the paper

To aid in electron transfer, these structures were picked such that they were highly planer \hl{(angle function?)}

@@

To be able to represent these molecular structures to the machine learning algorithms, a common naming scheme was formed such that it could be readily converted into a vector form. This naming scheme exploits the planar nature of benzobisazoles as well as their cruxiformic structure. Because of this regularity, they can be readily be structured in such a way that they can be converted into a format that is easy to represent to the machine learning algorithms. The naming scheme is broken down into three separate components. These parts are the core, the sides, and the options.

@@

The core is the central three fused ring system of the benzobisazoles. All cores are identified by a combination of three letters. The first letter indicates whether or not the structure is Cis or Trans which correspond to the letters C and T respectively. The second letter in the core name is the symbol for the element that is in the core without the $\pi$ bond (for consistency in naming this will always be the top right atom). This can be one of Oxygen (O), Sulfur (S), Nitrogen (N), Phosporous (P), or Carbon (C). The final letter is the remaining element's symbol which can be N, P, or C (for consistency this will always be the bottom right atom).

@@

Unexpectedly, the decay function with distance corrections did not perform any better than the regular decay feature vector. In theory, this should have worked better with heterogeneous side chains, but its use seemed to only increase the MAE.

There were a few anomalies in the results. The most noticeable is the long tail distribution of errors. A few molecules within the set had unusually high band gap errors (ex: $8A\_TON\_A\_A$, 1.1 eV). structures where features overlap ($A\_TON\_4ed\_A$, 0.5 eV)

To run the original DFT and TDDFT calculations for these molecules took about 35,0000 CPU hours. Calculating the same properties using these machine learning methods took about 4 seconds.

@@

For possible future work, the first thing to do would be to collect more data. For this study, the majority of the structures were monomers with 1-4 rings out from the core. This could be expanded to include larger polymers and monomers.

None of the feature vectors take into account the Kuhn expression for polymer chains. This leads to high errors when $n>1$ or $m>1$. The problem is not as pronounced when using SVM with Gaussian kernels because of how they decay \hl{(?)}. Preliminary work with this addition corrected seems to be the best way to account for polymerization of benzobisazoles.

Another possible expansion of work would be to test with more properties of the molecules such as total energy, dipole, and etc.

\hl{preliminary results have shown that the naive feature vector works the best with predicting total energy (117 hartree MAE with svm, 2800 hartree MAE with mean) (about 2x as effective). This could be because the total energy is more dependent on the structure as a whole rather than the localization to the core.}

As an end goal, it would be nice to generalize these machine learning methods such that they could be used for any structure. While the but perhaps there is a way to do it based off localized functional groups.

@@

\hl{In addition to the machine learning methods used in this study, there was also an attempt to use Neural Networks for the predictions. Unfortunately, the results were not satisfactory relative to the amount of time required (0.20 eV MAE). Drastic changes to the network made minimal differences to the final error. This is attributed to inexperience with proper methodologies with Neural Networks.}
\hl{Also, the methods mentioned in }\cite{hansen_assessment_2013} \hl{were tested and the results were not as good as any of the other name based features (0.25-0.35 eV). In addition to the accuracy problems, it had severe performance problems because it relys on $O(n^2)$ elements in each feature vector where $n$ is the number of atoms in the structure. To save time, only structures with <45 atoms total were tested, leaving about 450 structures.}

>>>>

For the side chains, structures were picked based on their ability to tune electronic properties. In the case of the main backbone structures, they were all picked because they are all highly planar and highly conjugated. Both of these features to aid in electron transfer. For the various substiuents, they picked based off their known ability to tune electronic properties \hl{(repeating...)}. The properties of this study have been shown in previous research to be correlated with Hammett Parameters. \hl{cool story...}

@@

Unfortunately, the machine learning algorithms used for this study do not natively understand molecular structures and instead only are able to function on vector forms. To rectify this, a common naming scheme was formed such that it could be readily converted into a vector form. This naming scheme exploits the planar nature of benzobisazoles as well as their cruxiformic structure. Because of this regularity, they can be readily be structured in such a way that they can be converted into a format that is easy to represent to the machine learning algorithms. The naming scheme is broken down into three separate components. These parts are the core, the sides, and the options.

@@

The core part of the structure is composed of a single three fused ring system. All cores are identified by a combination of three letters. The first letter indicates whether or not the structure is Cis or Trans which correspond to the letters C and T respectively. The second letter in the core name is the symbol for the element that is in the core without the $\pi$ bond (for consistency in naming this will always be the top right atom). This can be one of Oxygen (O), Sulfur (S), Nitrogen (N), Phosporous (P), or Carbon (C). The final letter is the remaining element's symbol which can be N, P, or C (for consistency this will always be the bottom right atom).

@@

Unexpectedly, the decay function with distance corrections did not perform any better than the regular decay feature vector. In theory, this should have worked better with heterogeneous side chains, but its use seemed to only increase the MAE. We think that the cause of this is due to the lack of sampling of largely heterogeneous side chains. So perhaps, if the same tests were done with only heterogeneous side chains adding the correction factor might decrease the total error.

There were a few anomalies in the results. The most noticeable is the long tail distribution of errors. As one might expect, structures where features overlap ($A\_TON\_4ed\_A$, 0.5 eV) there were higher MAE values. Unfortunately, a few molecules within the set had unusually high band gap errors (ex: $8A\_TON\_A\_A$, 1.1 eV) and we have no explanation for why that occurred.

To run the original DFT and TDDFT calculations for these molecules took about 35,0000 CPU hours. Calculating the same properties using these machine learning methods took about 4 seconds for each property for all 1100 structures. Because of the way that machine learning algorithms work, the majority of the time spent for each of these calculations is spent in the training phase; fortunately, this calculation is only required once so by . Once the models have been trained, these calculations are often reduced to just matrix multiplications leading to it taking just 0.1 seconds to calculate each property for all 1100 structures.

@@

For possible future work, the first thing to do would be to collect more data. For this study, the majority of the structures were monomers with 1-4 rings out from the core. This could be expanded to include larger polymers and monomers, especially with an emphasis on having more heterogeneous side chains. The functional groups could also be expanded to include things such as furan, pyrrole, or other conjugated systems.

None of the feature vectors take into account the Kuhn expression for polymer chains. Because many of these machine learning algorithms use linear models to fit, they do not accurately represent the nonlinear decay of influence seen in long polymers. This leads to high errors when $n>1$ or $m>1$. The problem is not as pronounced when using SVM with Gaussian kernels because of how they decay \hl{(?)}. \hl{Some preliminary work was done with predicting n=1,2,3,4 and then fitting the Kuhn expression for long polymers. This additional correction seems to be a good way to account for polymerization of benzobisazoles.}

% \begin{table}[H]
%   \centering
%   \caption{TON\_n20}
%   \begin{tabular}{lrrr}
%             Type      & HOMO (eV) & LUMO (eV) & Band Gap (eV) \\
%     \hline\hline
%             Predicted & -5.59 & -2.36 & 2.73 \\
%             Corrected & -5.99 & -3.36 & 2.13 \\
%             Actual    & -5.96 & -3.48 & 2.17 \\
%     \hline\hline
%   \end{tabular}
%   \label{tab:TONn20}
% \end{table}

% \begin{table}[H]
%   \centering
%   \caption{TON\_m20}
%   \begin{tabular}{lrrr}
%             Type      & HOMO (eV) & LUMO (eV) & Band Gap (eV) \\
%     \hline\hline
%             Predicted & -5.61 & -2.34 & 2.73 \\
%             Corrected & -5.84 & -2.20 & 3.10 \\
%             Actual    & -5.73 & -2.16 & 3.14 \\
%     \hline\hline
%   \end{tabular}
%   \label{tab:TONm20}
% \end{table}

Another possible expansion of work would be to test with more properties of the molecules such as total energy, dipole, and etc. \hl{preliminary results have shown that the naive feature vector works the best with predicting total energy (117 hartree MAE with svm, 2800 hartree MAE with mean) (about 2x as effective). This could be because the total energy is more dependent on the structure as a whole rather than the localization to the core.}

As an end goal, it would be nice to generalize these machine learning methods such that they could be used for any structure. \hl{While the} but perhaps there is a way to do it based off localized functional groups.

@@

\hl{In addition to the machine learning methods used in this study, there was also an attempt to use Neural Networks for the predictions. Unfortunately, the results were not satisfactory relative to the amount of time required (0.20 eV MAE). Drastic changes to the network made minimal differences to the final error. This is attributed to inexperience with proper methodologies with Neural Networks.}
\hl{Also, the methods mentioned in }\cite{hansen_assessment_2013} \hl{were tested and the results were not as good as any of the other name based features (0.25-0.35 eV). In addition to the accuracy problems, it had severe performance problems because it relys on $O(n^2)$ elements in each feature vector where $n$ is the number of atoms in the structure. To save time, only structures with $<45$ atoms total were tested, leaving about 450 structures.}
